### 2019-06-12 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Commentaires Christophe Ch6

% Stochastic bandits were historically introduced as a simple model for clinical trials, where arms correspond to some treatments with unknown efficacy \cite{Thompson33}.
% More recently, MAB models have been proven useful for different applications, like cognitive radio, where arms can model the vacancy of radio channels, as presented in Chapter~\ref{chapter:4}, or parameters of a dynamically configurable radio hardware \cite{Maghsudi16,Bonnefoi17,KerkoucheAlami18}.
% Another application is the design of recommender systems, where arms model the popularity of different items (\eg, news recommendation, \cite{Li10}).
%
For cognitive radio applications, as well as other applications such as recommender systems, the assumption that the arms distribution \emph{do not change over time} may be a big limitation.
Indeed, in cognitive radio and IoT networks, new devices can enter or leave the network, which impacts the availability of the radio channel they use to communicate.
And for instance in online recommendation, the popularity of items is also subject to trends.
Hence, there has been some interest on how to take those \emph{non-stationary} aspects into account within a multi-armed bandit model.

@@

It could seem that the problem is considered as solved by the Exp3.S algorithm, but it was actively studied since the seminal paper \cite{Auer02NonStochastic}.
Two main reasons explain the dynamic research on this problem:
first, it is well known that despite their qualities, exponential-weights algorithms like Exp3 can usually be greatly outperformed by UCB-based algorithms, for stochastic problems, and so it is expected that Exp3.S can be outperformed by other algorithms, thus any practical application might benefit from algorithms more efficient than Exp3.S.
%
The second reason is that passively adaptive algorithms function as black-box models: they do not try to detect changes and do not explain why they changed from one arm to another. A lot of the recent research has been focused on actively adaptive algorithms, precisely because they allow to interpret their decisions, by detecting changes on arms with statistical tests.
The piece-wise stationary MAB model was then studied by \cite{Kocsis06} and \cite{YuMannor09}.
%
This model is sometimes referred to as the abruptly changing \cite{WeiSrivastava18Abruptly}, or switching environments \cite{MellorShapiro13}.
%
In this model, described in full details in Section~\ref{sec:6:BanditSetting}, the (random) reward of arm $k$ at round $t$ has some mean $\mu_k(t)$, that is constant on intervals between two \emph{breakpoints}, and the regret is measured with respect to the \emph{current} best arm $k_t^* = \arg\max_i \mu_k(t)$.

@@

A \emph{piece-wise stationary bandit model} is characterized by a set of $K$ arms.
A (random) stream of rewards $(Y_{k,t})_{t\in\N^*}$ is associated to each arm $k \in [K]$. We assume that the rewards are bounded, and without loss of generality we assume that $Y_{k,t} \in [0,1]$.
We denote by $\mu_k(t) \eqdef  \bE[Y_{k,t}]$ the mean reward of arm $k$ at round $t$. At each round $t$, a decision maker has to select an arm $A(t)\in[K]$, based on past observation and receives the corresponding reward $r(t) = Y_{A(t),t}$.
At time $t$, we denote by $k_t^*$ an arm with maximal expected reward, \ie, $\mu_{k_t^*}(t) \eqdef \max_k \mu_k(t)$, called an optimal arm (possibly not unique).

@@

Note than when a breakpoint occurs, we do not assume that all the arms means  change, but that \emph{there exists} an arm whose mean has changed.
Depending on the application, many scenario can be meaningful: changes occurring on all arms simultaneously (due to some exogenous event), or only one or a few arms change at each breakpoint.
For instance, for a cognitive radio application in a IoT-like network, we can imagine that all the objects of one company use a fixed (or a subset of) channel(s), and on a particular day in a city, if the company deploys a lot of new objects, then one (or some) channel(s) see their mean occupancy rates abruptly change.

@@

We review previous works that studied the piece-wise stationary bandit model, or variants of this model.
To our knowledge, all the previous works are based on the idea of combining a classical bandit policy, such as Thompson sampling, \UCB{} or EXP3, with a strategy to account for changes in arms' distributions.
Following the vocabulary used in previous works, we make the distinction between passively and actively adaptive strategies.
Active strategies monitor the arms' rewards, by using change detection algorithms \cite{Basseville93}, and reset the history of pulls and rewards of one or all the arms as soon as a change is detected.
Passive strategies use a (fixed) discount factor or a limited memory size, while most active strategies use a growing memory.
% , while active strategies are usually tuned with a threshold for the statistical test.

@@

Both D-UCB and SW-UCB builds on a stationary policy, but for example EXP3.S from \cite{Auer02NonStochastic}
builds on the EXP3 policy, which is designed for the adversarial case.
EXP3.S actually achieves a good regret upper-bound, of the form $\cO(\sqrt{K \Upsilon_T T})$, with no additional prior knowledge except that of $T$ and $\Upsilon_T$.
It constitues a good baseline for our numerical experiments in Section~\ref{sec:6:NumericalExperiments},
even if we did find that EXP3.S performs worse that the most of the other approaches based on extensions of stationary bandit algorithms (\eg, SW-UCB).
% , so we do not include it in .
Similarly, previous works showed that older algorithms have no or weaker regret guarantees, and have been proven to be less efficient empirically, or are designed for more specific settings.

@@

\paragraph{Actively adaptive algorithms.}
%
We distinguish two families,
the first line of research uses frequentist change-point detectors \cite{Basseville93}, combined with stationary policies, usually index policies like \UCB.
When using an efficient CD algorithm with an efficient index policy, these approaches usually perform more efficiently than the passively algorithms.
%
The Adapt-EVE algorithm from \cite{Hartland06} uses a Page-Hikley (PH) Test and the \UCB{} policy, but no theoretical guarantee was given.
%
The Windowed-Mean Shift algorithm from \cite{YuMannor09} is more generic and combines any efficient bandit policy with a CD test based on a sliding window, but there approach is not applicable to the bandit setting of interest in this chapter, as they consider side observations.
%
The EXP3.R algorithm from \cite{Allesiardo15,Allesiardo17} combines a CD algorithm with EXP3, and the history of all arms are reset as soon as a suboptimal arm is detecting to become optimal.
A regret bound of $\cO(\Upsilon_T \sqrt{T \log(T)})$ was proven, even when $\Upsilon_T$ is known.

Two recent and related works use
the two-sided \CUSUM{} CD algorithm for \CUSUMUCB{} in \cite{LiuLeeShroff17}
or a specific and simpler CD algorithm for the Monitored UCB (\MUCB) algorithm introduced in \cite{CaoZhenKvetonXie18}.
%
The \CUSUM{} test is rather complicated and parametric: it uses the first $M$ samples (for a fixed $M$) from one arm to compute an average $\hat{u_0}$, and then builds two random walks, using the remaining observations for this arm. A change is detected when either random walks cross a threshold $h$.
It requires the tuning of three parameters, $M$ and $h$ as well as a drift correction parameter $\varepsilon\in(0,1)$.
The PH test, also studied in \cite{LiuLeeShroff17}, is similarly complex,
and we found that \PHT-\UCB{} perform very similarly to \CUSUMUCB{} in our experiments.
Even if it has the advantage of not requiring a parameter $M$,
it has no theoretical guarantee, so we do not include \PHT-\UCB{} in the experiments presented below.
%
In comparison, \MUCB{} uses a much simpler test, based on the $w$ most recent observations on one arm (for a fixed and even number $w$), and compares with a fixed threshold $h$ the difference of the sum of rewards for the first half and the second half. It is numerically much simpler, and has the advantage of using only a bounded memory (of the order $\bigO{K w}$ for $K$ arms).
%
Both approaches also introduced a mechanism to force a uniform exploration of each arm, parametered by $\alpha\in(0,1)$.
% \CUSUMUCB{} uses a randomized exploration, by sampling arm $k$ with probability $\alpha/K$ at each time step,
% and \MUCB{} uses a more deterministic scheme, to force exploring the $K$ arms (when the time since the last restart, $t - \tau$, is found to be in $[K]$ modulo $\left\lceil 1/\alpha \right\rceil$).
% Both solutions ensure that all arms are sampled enough on each stationary sequence, in order for the CD algorithm to have enough \iid{} samples to detect changes.

When all their parameters are tuned correctly, both approaches are proven to achieve a good regret upper-bound, of order $\cO(K \sqrt{\Upsilon_T T \log(K T)})$.
As both results are valid under different assumptions, we consider the two results to be the current state-of-the-art.
As for previous works, tuning their parameters requires to know both the horizon $T$ and the number of breakpoints $\Upsilon_T$, but most importantly it requires a prior knowledge of the problem difficulty, by assuming to know a lower bound on both the length of stationary segments (\eg, $L$ for \MUCB) and on the changes on the means of arms (\eg, $\varepsilon$ for \CUSUM).
%
\CUSUMUCB{} achieves a better regret upper bound, of the order of $\cO(\sqrt{\Upsilon_T T \log(T / \Upsilon_T)})$, but \emph{only} for Bernoulli distributions,
and when its $4$ parameters ($\alpha,\varepsilon,M,h$) are tuned based on problem-dependent knowledge.
%
\MUCB{} achieves a regret bounded by $\cO(\sqrt{\Upsilon_T T \log(T)})$, for bounded distributions, and when its $4$ parameters ($\alpha,w,b,\gamma$) are also tuned based on a problem-dependent knowledge of $\tilde{\delta}$ and $L$ (see Assumption~1 and Remark~1 in \cite{CaoZhenKvetonXie18}).

@@

Another interesting recent work is \cite{AuerGajaneOrtner18}, where the authors propose an efficient algorithm, AdSwitch, for the two-armed case ($K=2$).
It proceeds in episodes, starting a new episode when the algorithm detects a change in one of the arms. Each episode is an estimation phase, playing both arms alternatingly until their means can be distinguished, then an exploitation phase. With some low probability, an exploration phase is started that checks whether a change has occurred.
Similarly to what previous approaches did (\eg, \CUSUM{} uses a threshold $\varepsilon$), this phase checks for changes of a certain minimum magnitude, but the innovation in this work is that algorithm does not need any prior knowledge of the minimum magnitude.
Instead, it uses geometrically decreasing magnitudes $d_k = 2^{-i}$, meaning that even difficult changes should be detected, possibly in a later episode, as each episode only consider a few different values of $d_k$.
% Our  algorithmAdSwitch(shown  as  Algorithm  1)  proceeds  in  episodes,  where  a  newepisode starts when the algorithm detects a change in one of the arms.  Each episode startswith an estimation phase (Step 2), in which both arms are chosen alternatingly until theirmeans can be distinguished (Step 3).  After the estimation phase,  the algorithm exploitsthe empirical best arm (Step 5).  However,  in order to detect changes,  at each step withsome probability (Step 4) an exploration phase is started that checks whether a change hasoccurred (comparing the resulting empirical means to that obtained from the estimationphase).  During such exploration phases the algorithm checks for changes of certain magni-tudes (denoted bydi).  Note that our algorithm does not require knowledge of the numberof changesL.  The following theorem provides an upper bound on the expected regret ofour algorithm
They prove it achieves a regret bound of $\cO(\log(T) \sqrt{\Upsilon_T T})$, without a prior knowledge of $\Upsilon_T$.
However, this work as two drawbacks:
first, the algorithm currently only applies to two arms, and the generalization does not seem straightforward.
Then, the theoretical result is valid for ``sufficiently large constants $C_1$ and $C_2$'', with a large constant $C$ hiding in the $\cO(\bullet)$ notation.
We did some experiments with AdSwitch, even if it is not included in the benchmark presented below in Section~\ref{sec:6:NumericalExperiments}. It uses two parameters $C_1,C_2$ that are set at large values for the theoretical analysis,
and while we explored numerically different choices, it seemed that setting $C_1=C_2=1$ gave the best performance (which is outside of the comfort zone of the theoretical analysis). Even for this tuning, and for problems with just $K=2$ arms, empirically   this policy is performing poorly in comparison to other policies based on active CD detection and \klUCB{} indexes.
%  We did not find a practical tuning outperforming GLR-klUCB yet.
% It would be interesting to compare

\paragraph{Expert aggregation.}
%
Another of research on actively adaptive algorithms uses a Bayesian point of view.
A Bayesian CD algorithm is combined with Thompson Sampling in \cite{MellorShapiro13},
and more recently \cite{Alami17} introduced the Memory Bandit algorithm.
It is presented as efficient empirically, but no theoretical guarantee are given for these two works, and due to its complexity, we do not include it in our experiments.
The idea behind Memory Bandit is to use an expert aggregation algorithm, like Exp4 from \cite{Auer02} or our \Aggr{} algorithm from Section~\ref{sec:2:chooseYourPreferredBanditAlgorithm} \cite{Besson2018WCNC}, modified to efficiently aggregate an growing number of experts from techniques presented in \cite{Mourtada17}.
At each time step, a new expert is introduced, and experts corresponds to different Thompson sampling algorithms, each using a different history of pulls and rewards. Intuitively, after a change-point the newest experts will soon become the most efficient, as they are learning by using rewards drawn from the new distribution(s).
%
Note that the regret bound for these methods are still open-problems,
and they are computationally more costly,
therefore we choose to not include them in our experiments.

@@

This setting is quite different from the setting we are interested by in this chapter, and even if the most recent works on slowly-varying MAB models are very interesting, we preferred to focus on the piece-wise stationary (or abruptly changing) model,
hence we do no include any of these algorithms in the experiments presented in Section~\ref{sec:6:NumericalExperiments}.

\paragraph{Another point-of-view on prior knowledge of $\Upsilon_T$.}
%
In this chapter, like in our two main inspirational works \cite{CaoZhenKvetonXie18,LiuLeeShroff17},
we assume that the algorithm know in advance the number of break-points $\Upsilon_T$.
Another interesting point-of-view is the one presented by \cite{WeiSrivastava18Abruptly,WeiSrivastava18Distributed},
where the authors assume that there exists a number $\nu\in(0,1)$ for which $\Upsilon_T = \smallO{T^{\nu}}$,
and they also assume that any algorithm tackling such problems can know in advance the value of $\nu$ (or an upper-bound on $\nu$), and tweak its parameters from this value $\nu$, but cannot know the exact value of $\Upsilon_T$.
%
This interpretation is interesting too,
% but due to time and space constraints, we preferred to not explore it further.
but empirically we found that the SW-UCB\# algorithm proposed in \cite{WeiSrivastava18Abruptly} performs comparably to SW-UCB, and thus we did not include it in the experiments presented below in Section~\ref{sec:6:NumericalExperiments}.
%
% \TODOL{Speak about the article \cite{WeiSrivastava18Abruptly,WeiSrivastava18Distributed} and their point of view on $\Upsilon_T = \bigO{T^{\nu}}$ with known $\nu$ but unknown $\Upsilon_T$}

\paragraph{Adversarial or non-stationary contextual bandits.}
%
For contextual bandits, two very recent works are \cite{Luo18} and \cite{ChenLeeLuoWei2019}.
While both works target a more general setting, their algorithms are also applicable to non-contextual bandits, \ie, classical bandits like the model in this chapter.
From Table~1 in \cite{Luo18}, one can see that their two algorithms Exp4.S and Ada-ILTCB recover the same regret guarantees as we do, in the non-contextual case:
that is, a $\cO(\Upsilon_T \sqrt{T}$ regret without knowing the number of changes and $\cO(\sqrt{\Upsilon_T T})$ with this knowledge.
We discuss both these algorithms:
%
first, Exp4.S actually reduces to Exp3.S in the non-contextual case, and Exp3.S is discussed above.
% As already acknowledged in the answer to  Reviewer 1,  we were indeed unaware of the very good theoretical properties of this algorithm in the piece-wise non-stationary setting. We will revise our state-of-the-art section and include the Exp3.S algorithm as a baseline in our experiments. However, preliminary experiments reveal that this algorithm is much less efficient that GLR-klUCB.
%
Then, the Ada-ILTCB algorithm is however a less appealing candidate for our setting, for the following reason.
This algorithm requires as input a parameter $L$ which needs to be an upper bound on the largest stationary sequence in the problem, in order for its regret to scale as $\Upsilon_T \sqrt{L}$ (neglecting $K$ and $\log$ factors).
It is claimed that the tuning $L=T/\Upsilon_T$ yields the optimal bound $\sqrt{\Upsilon_T T}$.
However, this tuning is only possible in a ``balanced'' instance without stationary sequence longer than $T/\Upsilon_T$.
In a piece-wise stationary instance with a stationary sequence of length $T/2$, one cannot get better than $\cO(\Upsilon_T \sqrt{T})$ regret, by setting $L=T/2$ (see for instance the problem 4 in our benchmark below in Section~\ref{sec:6:NumericalExperiments}), and as such the regret bound of Ada-ILTCB appear much worse than the announced $\cO(\sqrt{\Upsilon_T T})$ bound.
%
Besides, no experiments are reported by \cite{Luo18}, and the Ada-ILTCB algorithm seems much more complicated to implement than other alternatives,
so we prefer to not include it in the experiments presented later.
% with a valid tuning of L in each case.

@@

Assume that we collect independent samples $X_1,X_2,\dots$ all from some distributions supported in $[0,1]$.
We want to discriminate between two possible scenarios: all the samples come from distributions that have a common mean $\mu_0$, or there exists a \emph{change-point} $\tau > 1$ such that $X_1,\dots,X_\tau$ have some mean $\mu_0$ and $X_{\tau +1},X_{\tau+2},\dots$ have a different mean $\mu_1 \neq \mu_0$.
%
A sequential change-point detector is a \emph{stopping time}\footnote{~Stopping times are for instance presented formally in Chapter~3 of \cite{LattimoreBanditAlgorithmsBook}.} $\widehat{\tau}$ with respect to the filtration $\cF_t = \sigma(X_1,\dots,X_t)$ such that $(\hat \tau < \infty)$ means that we reject the hypothesis
$\cH_0 : \left(\exists \mu_0 \in [0,1]: \forall t\in\N^*, \bE[X_t] = \mu_0\right)$.

@@

This section give the finite-time regret bounds for the two variants, and interpretations of the different results.
Although we only prove one bound, the other proof can be found in \cite{Besson2019GLRT}.

% ----------------------------------------------------------------------------
\subsection{Results for \GLRklUCB{} using Global Changes}

@@

\subsection{Proof for \GLRklUCB{} with global changes}
\label{proof:6:mainRegretBoundGlobal}

>>>>

% Stochastic bandits were historically introduced as a simple model for clinical trials, where arms correspond to some treatments with unknown efficacy \cite{Thompson33}.
% More recently, MAB models have been proven useful for different applications, like cognitive radio, where arms can model the vacancy of radio channels, as presented in Chapter~\ref{chapter:4}, or parameters of a dynamically configurable radio hardware \cite{Maghsudi16,Bonnefoi17,KerkoucheAlami18}.
% Another application is the design of recommender systems, where arms model the popularity of different items (\eg, news recommendation, \cite{Li10}).
%
For cognitive radio applications, as well as other applications such as recommender systems, the assumption that the arms distribution \emph{do not change over time} may be a big limitation.
Indeed, in cognitive radio or IoT networks, new devices can enter or leave the network, which impacts the availability of the radio channel they use to communicate, making it non stationary.
And for instance in online recommendation, the popularity of items is also subject to trends.
Hence, there has been some interest on how to take those \emph{non-stationary} aspects into account within a multi-armed bandit model.

@@

It could seem that the problem is considered as solved by the Exp3.S algorithm, but it was actively studied since the seminal paper \cite{Auer02NonStochastic}.
Two main reasons explain the dynamic research on this problem:
first, it is well known that despite their qualities, exponential-weights algorithms like Exp3 can usually be greatly outperformed by UCB-based algorithms, for stochastic problems, and so it is expected that Exp3.S can be outperformed by other algorithms, thus any practical application might benefit from algorithms more efficient than Exp3.S.
%
The second reason is that passively adaptive algorithms function as black-box models: they do not try to detect changes and do not explain why they changed from one arm to another. A lot of the recent research has been focused on actively adaptive algorithms, precisely because they allow to interpret their decisions, by detecting changes on arms with statistical tests.
The piece-wise stationary MAB model was then studied by \cite{Kocsis06} and \cite{YuMannor09}.
%
This model is sometimes referred to as the abruptly changing \cite{WeiSrivastava18Abruptly}, or switching environments \cite{MellorShapiro13}.
%
In this model, described in full details in Section~\ref{sec:6:BanditSetting}, the (random) reward of arm $k$ at round $t$ has some mean $\mu_k(t)$, that is constant on intervals between two \emph{breakpoints}, and the regret is measured with respect to the \emph{current} best arm $k_t^* \eqdef \arg\max_i \mu_k(t)$.

@@

A \emph{piece-wise stationary bandit model} is characterized by a set of $K$ arms.
A (random) stream of rewards $(Y_{k,t})_{t\in\N^*}$ is associated to each arm $k \in [K]$. We assume that the rewards are bounded, and without loss of generality we assume that $Y_{k,t} \in [0,1]$.
We denote by $\mu_k(t) \eqdef  \bE[Y_{k,t}]$ the mean reward of arm $k$ at round $t$. At each round $t$, a decision maker has to select an arm $A(t)\in[K]$, based on past observation and receives the corresponding reward $r(t) \eqdef Y_{A(t),t}$.
At time $t$, we denote by $k_t^*$ an arm with maximal expected reward, \ie, $\mu_{k_t^*}(t) \eqdef \max_k \mu_k(t)$, called an optimal arm (possibly not unique).

@@

Note than when a breakpoint occurs, we do not assume that all the arms means  change, but that \emph{there exists} an arm whose mean has changed.
Depending on the application, many scenario can be meaningful: changes occurring on all arms simultaneously (due to some exogenous event), or only one or a few arms change at each breakpoint.
For instance, for a cognitive radio application in a IoT-like network, we can imagine that all the objects of one company (or network provider) use a fixed (or a subset of) channel(s), and on a particular day in a city, if the company deploys a lot of new objects, then one (or some) channel(s) see their mean occupancy rates abruptly change.

@@

We review previous works that studied the piece-wise stationary bandit model, or variants of this model.
To the best of our knowledge, all the previous works are based on the idea of combining a classical bandit policy, such as Thompson sampling, \UCB{} or EXP3, with a strategy to account for changes in arms' distributions.
Following the vocabulary used in previous works, we make the distinction between passively and actively adaptive strategies.
Active strategies monitor the arms' rewards, by using change detection algorithms \cite{Basseville93}, and reset the history of pulls and rewards of one or all the arms as soon as a change is detected.
Passive strategies use a (fixed) discount factor or a limited memory size, while most active strategies use a growing memory.
% , while active strategies are usually tuned with a threshold for the statistical test.

@@

Both D-UCB and SW-UCB build on a stationary policy, but for example EXP3.S from \cite{Auer02NonStochastic}
builds on the EXP3 policy, which is designed for the adversarial case.
EXP3.S actually achieves a good regret upper-bound, of the form $\cO(\sqrt{K \Upsilon_T T})$, with no additional prior knowledge except that of $T$ and $\Upsilon_T$.
It constitues a good baseline for our numerical experiments in Section~\ref{sec:6:NumericalExperiments},
even if we did find that EXP3.S performs worse that the most of the other approaches based on extensions of stationary bandit algorithms (\eg, SW-UCB).
% , so we do not include it in .
Similarly, previous works showed that older algorithms have no or weaker regret guarantees, and have been proven to be less efficient empirically, or are designed for more specific settings.

@@

\paragraph{Actively adaptive algorithms.}
%
We distinguish two families,
the first line of research uses frequentist change-point detectors \cite{Basseville93}, combined with stationary policies, usually index policies like \UCB.
When using an efficient Change-point Detector (CD) algorithm with an efficient index policy, these approaches usually perform more efficiently than the passively algorithms.
%
The Adapt-EVE algorithm from \cite{Hartland06} uses a Page-Hikley (PH) Test and the \UCB{} policy, but no theoretical guarantee was given.
%
The Windowed-Mean Shift algorithm from \cite{YuMannor09} is more generic and combines any efficient bandit policy with a CD test based on a sliding window, but their approach is not applicable to the bandit setting of interest in this chapter, as they consider side observations.
%
The EXP3.R algorithm from \cite{Allesiardo15,Allesiardo17} combines a CD algorithm with EXP3, and the history of all arms are reset as soon as a suboptimal arm is detecting to become optimal.
A regret bound of $\cO(\Upsilon_T \sqrt{T \log(T)})$ was proven, even when $\Upsilon_T$ is known.

Two recent and related works use
the two-sided \CUSUM{} CD algorithm for \CUSUMUCB{} in \cite{LiuLeeShroff17}
or a specific and simpler CD algorithm for the Monitored UCB (\MUCB) algorithm introduced in \cite{CaoZhenKvetonXie18}.
%
The \CUSUM{} test is rather complicated and parametric: it uses the first $M$ samples (for a fixed $M$) from one arm to compute an average $\hat{u_0}$, and then builds two random walks, using the remaining observations for this arm. A change is detected when either random walks cross a threshold $h$.
It requires the tuning of three parameters, $M$ and $h$ as well as a drift correction parameter $\varepsilon\in(0,1)$.
The PH test, also studied in \cite{LiuLeeShroff17}, is similarly complex,
and we found that \PHT-\UCB{} perform very similarly to \CUSUMUCB{} in our experiments.
Even if it has the advantage of not requiring a parameter $M$,
it has no theoretical guarantee, so we do not include \PHT-\UCB{} in the experiments presented below.
%
In comparison, \MUCB{} uses a much simpler test, based on the $w$ most recent observations on one arm (for a fixed and even number $w$), and compares with a fixed threshold $h$ the difference of the sum of rewards for the first half and the second half. It is numerically much simpler, and has the advantage of using only a bounded memory (of the order $\bigO{K w}$ for $K$ arms).
%
Both approaches also introduced a mechanism to force a uniform exploration of each arm, parametered by $\omega\in(0,1)$.
% \CUSUMUCB{} uses a randomized exploration, by sampling arm $k$ with probability $\omega/K$ at each time step,
% and \MUCB{} uses a more deterministic scheme, to force exploring the $K$ arms (when the time since the last restart, $t - \tau$, is found to be in $[K]$ modulo $\left\lceil 1/\omega \right\rceil$).
% Both solutions ensure that all arms are sampled enough on each stationary sequence, in order for the CD algorithm to have enough \iid{} samples to detect changes.

When all their parameters are tuned correctly, both approaches are proven to achieve a good regret upper-bound, of order $\cO(K \sqrt{\Upsilon_T T \log(K T)})$.
As both results are valid under different assumptions, we consider the two results to be the current state-of-the-art.
As for previous works, tuning their parameters requires to know both the horizon $T$ and the number of breakpoints $\Upsilon_T$.
But most importantly it requires a prior knowledge of the problem difficulty, by assuming to know a lower bound on both the length of stationary segments (\eg, $L$ for \MUCB) and on the changes on the means of arms (\eg, $\varepsilon$ for \CUSUM).
%
\CUSUMUCB{} achieves a better regret upper bound, of the order of $\cO(\sqrt{\Upsilon_T T \log(T / \Upsilon_T)})$, but \emph{only} for Bernoulli distributions,
and when its $4$ parameters ($\omega,\varepsilon,M,h$) are tuned based on problem-dependent knowledge.
%
\MUCB{} achieves a regret bounded by $\cO(\sqrt{\Upsilon_T T \log(T)})$, for bounded distributions, and when its $4$ parameters ($\omega,w,b,\gamma$) are also tuned based on a problem-dependent knowledge of $\tilde{\delta}$ and $L$ (see Assumption~1 and Remark~1 in \cite{CaoZhenKvetonXie18}).

@@

Another interesting recent work is \cite{AuerGajaneOrtner18}, where the authors propose an efficient algorithm, AdSwitch, for the two-armed case ($K=2$).
It proceeds in episodes, starting a new episode when the algorithm detects a change in one of the arms.
Each episode consists of three phases, estimation, exploitation then exploration.
First, an estimation phase where both arms are played alternatingly until their means can be distinguished.
Then follows an exploitation phase, and finally with some low probability, an exploration phase is started that checks whether a change has occurred.
Similarly to what previous approaches did (\eg, \CUSUM{} uses a threshold $\varepsilon$), this phase checks for changes of a certain minimum magnitude, but the innovation in this work is that this algorithm does not need any prior knowledge of the minimum magnitude.
Instead, it uses geometrically decreasing magnitudes $d_k = 2^{-i}$, meaning that even difficult changes should be detected, possibly in a later episode, as each episode only considers a few different values of $d_k$.
% Our  algorithmAdSwitch(shown  as  Algorithm  1)  proceeds  in  episodes,  where  a  newepisode starts when the algorithm detects a change in one of the arms.  Each episode startswith an estimation phase (Step 2), in which both arms are chosen alternatingly until theirmeans can be distinguished (Step 3).  After the estimation phase,  the algorithm exploitsthe empirical best arm (Step 5).  However,  in order to detect changes,  at each step withsome probability (Step 4) an exploration phase is started that checks whether a change hasoccurred (comparing the resulting empirical means to that obtained from the estimationphase).  During such exploration phases the algorithm checks for changes of certain magni-tudes (denoted bydi).  Note that our algorithm does not require knowledge of the numberof changesL.  The following theorem provides an upper bound on the expected regret ofour algorithm
They prove it achieves a regret bound of $\cO(\log(T) \sqrt{\Upsilon_T T})$, without a prior knowledge of $\Upsilon_T$.
However, this work as two drawbacks:
first, the algorithm currently only applies to two arms, and the generalization does not seem straightforward.
Then, the theoretical result is valid for ``sufficiently large constants $C_1$ and $C_2$'', with a large constant $C$ hiding in the $\cO(\bullet)$ notation.
We did some experiments with AdSwitch, even if it is not included in the benchmark presented below in Section~\ref{sec:6:NumericalExperiments}. It uses two parameters $C_1,C_2$ that are set at large values for the theoretical analysis,
and while we explored numerically different choices, it seemed that setting $C_1=C_2=1$ gave the best performance (which is outside of the comfort zone of the theoretical analysis). Even for this tuning, and for problems with just $K=2$ arms, empirically   this policy is performing poorly in comparison to other policies based on active CD detection and \klUCB{} indexes.
%  We did not find a practical tuning outperforming GLR-klUCB yet.
% It would be interesting to compare

\paragraph{Expert aggregation.}
%
Another of research on actively adaptive algorithms uses a Bayesian point of view.
A Bayesian CD algorithm is combined with Thompson Sampling in \cite{MellorShapiro13},
and more recently \cite{Alami17} introduced the Memory Bandit algorithm.
It is presented as efficient empirically, but no theoretical guarantee are given for these two works, and due to its complexity, we do not include it in our experiments.
The idea behind Memory Bandit is to use an expert aggregation algorithm, like Exp4 from \cite{Auer02} or our \Aggr{} algorithm from Section~\ref{sec:2:chooseYourPreferredBanditAlgorithm} \cite{Besson2018WCNC}, modified to efficiently aggregate a growing number of experts from techniques presented in \cite{Mourtada17}.
At each time step, a new expert is introduced, and experts correspond to different Thompson sampling algorithms, each using a different history of pulls and rewards. Intuitively, after a change-point the newest experts will soon become the most efficient, as they are learning by using rewards drawn from the new distribution(s).
%
Note that the regret bound for these methods are still open-problems,
and they are computationally more costly,
therefore we choose to not include them in our experiments.

@@

This setting is quite different from the setting we are considering in this chapter, and even if the most recent works on slowly-varying MAB models are very interesting, we preferred to focus on the piece-wise stationary (or abruptly changing) model,
hence we do no include any of these algorithms in the experiments presented in Section~\ref{sec:6:NumericalExperiments}.

\paragraph{Another point-of-view on prior knowledge of $\Upsilon_T$.}
%
In this chapter, like in our two main inspirational works \cite{CaoZhenKvetonXie18,LiuLeeShroff17},
we assume that the algorithm knows in advance the number of break-points $\Upsilon_T$.
Another interesting point-of-view is the one presented by \cite{WeiSrivastava18Abruptly,WeiSrivastava18Distributed},
where the authors assume that there exists a number $\nu\in(0,1)$ for which $\Upsilon_T = \smallO{T^{\nu}}$,
and they also assume that any algorithm tackling such problems can know in advance the value of $\nu$ (or an upper-bound on $\nu$), and tweak its parameters from this value $\nu$, but cannot know the exact value of $\Upsilon_T$.
%
This interpretation is interesting too,
% but due to time and space constraints, we preferred to not explore it further.
but empirically we found that the SW-UCB\# algorithm proposed in \cite{WeiSrivastava18Abruptly} performs comparably to SW-UCB, and thus we did not include it in the experiments presented below in Section~\ref{sec:6:NumericalExperiments}.
%
% \TODOL{Speak about the article \cite{WeiSrivastava18Abruptly,WeiSrivastava18Distributed} and their point of view on $\Upsilon_T = \bigO{T^{\nu}}$ with known $\nu$ but unknown $\Upsilon_T$}

\paragraph{Adversarial or non-stationary contextual bandits.}
%
For contextual bandits, two very recent works are \cite{Luo18} and \cite{ChenLeeLuoWei2019}.
While both works target a more general setting, their algorithms are also applicable to non-contextual bandits, \ie, classical bandits like the model in this chapter.
From Table~1 in \cite{Luo18}, one can see that their two algorithms Exp4.S and Ada-ILTCB recover the same regret guarantees as we do, in the non-contextual case:
that is, a $\cO(\Upsilon_T \sqrt{T})$ regret without knowing the number of changes and $\cO(\sqrt{\Upsilon_T T})$ with this knowledge.
We discuss both these algorithms:
%
first, Exp4.S actually reduces to Exp3.S in the non-contextual case, and Exp3.S is discussed above.
% As already acknowledged in the answer to  Reviewer 1,  we were indeed unaware of the very good theoretical properties of this algorithm in the piece-wise non-stationary setting. We will revise our state-of-the-art section and include the Exp3.S algorithm as a baseline in our experiments. However, preliminary experiments reveal that this algorithm is much less efficient that GLR-klUCB.
%
Then, the Ada-ILTCB algorithm is however a less appealing candidate for our setting, for the following reason.
This algorithm requires as input a parameter $L$ which needs to be an upper bound on the largest stationary sequence in the problem, in order for its regret to scale as $\Upsilon_T \sqrt{L}$ (neglecting $K$ and $\log$ factors).
It is claimed that the tuning $L=T/\Upsilon_T$ yields the optimal bound $\sqrt{\Upsilon_T T}$.
However, this tuning is only possible in a ``balanced'' instance without stationary sequence longer than $T/\Upsilon_T$.
In a piece-wise stationary instance with a stationary sequence of length $T/2$, one cannot get better than $\cO(\Upsilon_T \sqrt{T})$ regret, by setting $L=T/2$ (see for instance the problem 4 in our benchmark below in Section~\ref{sec:6:NumericalExperiments}), and as such the regret bound of Ada-ILTCB appears much worse than the announced $\cO(\sqrt{\Upsilon_T T})$ bound.
%
Besides, no experiments are reported by \cite{Luo18}, and the Ada-ILTCB algorithm seems much more complicated to implement than other alternatives,
so we prefer to not include it in the experiments presented later.
% with a valid tuning of L in each case.

@@

Assume that we collect independent samples $X_1,X_2,\dots$ all from some distributions supported in $[0,1]$.
We want to discriminate between two possible scenarios: all the samples come from distributions that have a common mean $\mu_0$, or there exists a \emph{change-point} $\tau > 1$ such that $X_1,\dots,X_\tau$ have some mean $\mu_0$ and $X_{\tau +1},X_{\tau+2},\dots$ have a different mean $\mu_1 \neq \mu_0$.
%
A sequential change-point detector is a \emph{stopping time}\footnote{~Stopping times are for instance presented formally in Chapter~3 of \cite{LattimoreBanditAlgorithmsBook}.} $\widehat{\tau}$ with respect to the filtration $\cF_t \eqdef \sigma(X_1,\dots,X_t)$ such that $(\hat \tau < \infty)$ means that we reject the hypothesis
$\cH_0 : \left(\exists \mu_0 \in [0,1]: \forall t\in\N^*, \bE[X_t] = \mu_0\right)$.

@@

This section gives the finite-time regret bounds for the two variants, and interpretations of the different results.
Although we only prove one bound, the other proof can be found in \cite{Besson2019GLRT}.

% ----------------------------------------------------------------------------
\subsection{Results for \GLRklUCB{} using \emph{global restarts}}

@@

\subsection{Proof for \GLRklUCB{} with \emph{global restarts}}
\label{proof:6:mainRegretBoundGlobal}
